### 2017-07-11 introduction.tex
Minor

\section{Introduction}

% Describe data science and a lab

While science has long been collaborative and data driven, today, more than ever before, 
researchers are dependent on technologies that enable their ability to safely store, 
effectively manage, and quickly analyze vast amounts of
data. Moreover, individual datasets that are of interest to science and scholarship 
vary widely in size, often take on heterogeneous structures, and command varying 
levels of sensitivity. In fact, data sensitivity is often `custom' -defined specifically by
data use agreements that can be so granular as to vary from user to user and institution 
to institution. At the very same time, however, a typical research setting involves 
collaborators from multiple institutions that often share data, analysis techniques, 
and results. This is, in fact, a cornerstone of modern research. Yet, when data are 
particularly sensitive and their use agreements limit access and activities at the 
level of individual users, research collaboration can either lead to inadvertent 
breaches of agreement or can grind to a halt under administrative burden.

% Describe data and security concerns

Data sensitivity and security concerns are well founded. While many datasets are public the most
valuable ones are often considered private due to legal or contractual obligations under which the data
is shared, or even governmental regulations due to the nature of the data. The limited and proprietary nature
of the data further increases the perceived value to the researcher. However this also adds the burden of
ensuring the security of the data to the researchers.

% Ability to compute effectively and securely.
To address these concerns, a myriad of computational infrastructures
have been developed. These infrastructures, including our own \NAMENS, 
implement the software required to securely store and analyze protected
data. However, in almost every case, there is one significant weakness in these systems
and that is the humans tasked with administering and managing 
data collections. Most often, the policies and processes for managing
the data are ad hoc and lack software support.

To address these challenges we introduce two abstractions: 
the safe-collection to define the scope of a data collection
and the associated policies and the role of a data steward
to manage safe collections.  These two constructs allow us 
to define, manage, and audit access to, and use of, of 
sensitive data.  We implement these constructs in the

%Here are some critical problems researchers face in this space:
%
%\begin{itemize}
%\item Datasets are large, posing problems for storage, compute and cost.
%\item Varying degrees of sensitivity in data-sets require configurable security
%\item Share access, analyses across individuals and groups.
%\end{itemize}