### 2017-07-14 relatedwork.tex
Adding stuff

\section{Related Work}

Secure data enclaves have a long and varied history starting with Air gapped systems that are physically
disconnected from the internet
NORC Data Enclave