### 2017-07-15 introduction.tex
edits to intro

\section{Introduction}

% Describe data science and a lab

While science has long been collaborative and data driven, today, more than ever before,
researchers are dependent on technologies that enable their ability to safely store,
effectively manage, and quickly analyze vast amounts of
data. Moreover, individual datasets that are of interest to science and scholarship
vary widely in size, often take on heterogeneous structures, and command varying
levels of sensitivity. In fact, data sensitivity is often `custom' -defined specifically by
data use agreements that can be so granular as to vary from user to user and institution
to institution. At the very same time, however, a typical research setting involves
collaborators from multiple institutions that often share data, analysis techniques,
and results. This is, in fact, a cornerstone of modern research. Yet, when data are
particularly sensitive and their use agreements limit access and activities at the
level of individual users, research collaboration can either lead to inadvertent
breaches of agreement or can grind to a halt under administrative friction.

% Describe data and security concerns

Data sensitivity and security concerns are well founded. While many datasets are public, the most
valuable ones are often considered private due to legal or contractual obligations under which the data
is shared, or even governmental restriction. The limited and proprietary nature
of particular datasets can further increase perceived value to researchers given that the space of possible questions and possible answers that can be derived from the data is, in a sense, hidden from competition. At the same time, sensitivity adds the burden of
ensuring the security of the data to the researchers.

% Ability to compute effectively and securely.
To address these concerns, myriad computational infrastructures
have been developed. These infrastructures, including \NAMENS \cite{babuji2016cloud},
implement the software required to securely store and analyze protected
data. However, in almost every case, there is one significant weakness to these systems: the individual humans tasked with administering and managing
data collections are fallible and prone to error. Most often, the policies and processes for managing
the data are ad hoc and lack software support.

To address these challenges we introduce two abstractions:
the `safe-collection' to define the scope of a data collection
and the associated policies and the role of a data `steward'
to manage safe collections. These two constructs allow us
to define, manage, and audit access to, and use of
sensitive data.  We implement these constructs in \NAME and
describe how such implementation enhances security without compromising collaborative agility.

%Here are some critical problems researchers face in this space:
%
%\begin{itemize}
%\item Datasets are large, posing problems for storage, compute and cost.
%\item Varying degrees of sensitivity in data-sets require configurable security
%\item Share access, analyses across individuals and groups.
%\end{itemize}