### 2017-07-26 summary.tex
minor to summary

There are significant opportunities for new scientific discoveries to be made
through the collaborative analysis of sensitive and protected
data. While many approaches to storing such data
have been adopted in production deployments, in most cases there are
considerable barriers that overly restrict data access and analysis. 
Importantly, there is little functionality that enales dataset owners and administrators
to define and apply sophisticated administrative models and policies 
for securely managing that data.

To address these concerns we have developed two important abstractions
that allow data owners to securely manage data in ways that permit
controlled access and analysis. The `secure collection' provides a 
policy-based approach for defining possible interactions with a particular
dataset. `Data stewards' are associated with collections to then
manage the collections and review the actions of their users. These
constructs have been implemented in the \NAME data enclave
and have since been adopted to protect a range of sensitive
social science data. Initial experiences with the model
have been positive, with data owners valuing the ability 
to securely manage collections, and users valuing the clearly
defined policies for data access and analysis.